### 2014-04-07 final_presentation/presentation.tex
Added blueprint files for presentation and paper.

\documentclass[10pt]{beamer}

\usepackage{gensymb}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{attrib}
\usepackage{textpos}

\usetheme{Dresden}
\usecolortheme{dove}
\usenavigationsymbolstemplate{}
\title{Research?}
\author{Pieter Brederode, Timo Koppenberg, \\ Jelco Bodewes, Anton Golov}
\institute{B3OMI}

\begin{document}

\begin{frame}
    \maketitle
\end{frame}

\section{Introduction}
\begin{frame}{Overview}
    \begin{itemize}
        \item We did research!
        \item That's it, really.
    \end{itemize}
\end{frame}

\section{Conclusions}

\begin{frame}{Results}
    \begin{itemize}
        \item What results?
        \item What are you talking about?
    \end{itemize}
\end{frame}

\end{document}