### 2014-04-17 paper/paper.tex
Tried to make things more formal.

\documentclass[12pt,a4paper]{article}
% vim: set textwidth=100:

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{hyperref}

\title{Associative Data Structures for Website Filtering}
\author{Jelco Bodewes, Pieter Brederode \\ Anton Golov, Timo Koppenberg }

\begin{document}
    \maketitle

    \begin{abstract}
        We have researched the performance of the hash table, binary search tree, and skip list
        associative datastructures in the context of a web filter.  Our measurements were made with
        structures of various sizes (ranging from $2^8$ up to $2^{20}$) and with different possible
        key types (from 4 to 120 bytes).  In this paper, we describe our implementation and
        empirically show that given a sufficiently high number of entries, a hash table has the
        highest insertion and query speeds, while the skip list requires the least memory per
        element.
        % And when I say "hash table", I mean libstdc++'s unordered_map, and when I say bst I mean
        % libstdc++'s map, and if you took different implementations you'd probably get different
        % results...
    \end{abstract}

    \section{Introduction}
    A website filter is a program that categorises websites in order to restrict access to designated
    sites a user might visit. Commonly, the filter operates by keeping a blacklist of sites a user may not visit
    unless they have certain credentials, or alternatively, keeping a whitelist and prohibiting any
    sites not on it.  As the world-wide web is growing, these lists have grown larger to a point where
    the speed of the underlying data structure has become a key issue.  The time of both inserting and
    looking up websites must be kept minimal. Moreover, the increased number of entries make memory
    usage an equally significant concern.

    In technical terms, we are interested in the best data structure for associating many URLs with
    their respective permissions, be it whitelisted or blacklisted for certain users.  We have
    chosen to compare the performance of the binary search tree, the hash map, and the skip list, as
    these are commonly employed in practice.  It is of particular interest how each data structure
    will scale as the number of elements rises.  We have thus chosen for our main research question
    the following:
    % Guys, did you notice we did absolutely nothing about predicting this?

    \begin{quotation}
        How does the number of entries influence the performance of various associative data
        structures?
    \end{quotation}

    While the decision of data structure might mostly be dependant on the number of entries, the
    complexity of the key might also be of concern. Where permissions per page would be preferred, a
    more resource optimal solution might be found in filtering only on domain or even IP adress.  In
    researching the influence of this we formulate a secondary research question:

    \begin{quotation}
        How does the complexity of the key influence the preformance of various associative data structures?
    \end{quotation}

    In this paper, we attempt to answer these questions. Section~\ref{sec:plan} provides details
    about our research plan and the scopes of our research. Section~\ref{sec:implementation} is
    about the implementation of our setup in C++.  The rest of the paper states our results and the
    conclusions that can be drawn from them.

    \section{Plan}
    \label{sec:plan}

    \section{Implementation}
    \label{sec:implementation}

    We used C++ implementations of the data structures involved, and also the language for our
    measuring code.  We chose C++ as it is an industry standard\footnote{ref}, provides the
    necessary tools, and does not carry the risk of garbage collection pauses or effects due to
    ``warming-up time''.

    \subsection*{Data Structures}

    Using C++ as the implementation language, we chose libstdc++'s implementations of the standard class
    templates \texttt{std::map} and \texttt{std::unordered\_map} for the binary search
    tree and the hash table respectively.  As the C++ standard library does not contain a skip
    list\footnote{reference to the standard here}, we used an open-source third party library,
    \texttt{CSSkipList}\footnote{reference to library home page here}.  We also used libstdc++'s
    implementation of \texttt{std::string} for our string handling\footnote{specify?}needs.

    \subsection*{Data}
    For our data, we took the system's dictionary file and the list of top level domains from Wikipedia
    and combined these to generate a list of items in the form \texttt{www.\$word.\$tld}, where
    \texttt{\$word} is one of words from the the dictionary and \texttt{\$tld} is a top level domain.  Only
    words consisting of the lowercase letters \texttt{a} through \texttt{z} were used.  Constraining the
    words to consist of 8 to 12 characters and limiting the top-level domains to exactly two letters, 
    all possible combinations were generated.  For the full paths, a number of domains were selected
    and each was appended a random alphanumeric string of eighty characters.  The exact implementation
    can be viewed in the sourcecode\footnote{ref to apendix or code dump}.

    \subsection*{Measurements}
    For our time measurements, we used libstdc++'s implementation of the standard
    \texttt{std::chrono::high\_resolution\_clock} class.

    Memory measurements were performed by providing a custom allocator which tracked the number of memory
    requests.  Note that the latter deviates from our initial plan, where we stated\footnote{ref?}we would
    use the POSIX API\footnote{ref to the posix api}.  Unfortunately, that turned out to be too
    coarse-grained for our purposes and we could thus get no accurate measurements.  The
    allocator-based technique we employed did require some slight modifications to the skip list
    implementation we used, but this should not have caused a significant slowdown. We assume that any 
    influence will only have made the tests more fair, as now allocation was done the same way in all 
    three setups.

    After some tests we chose to decrease the number of runs used for measurements.  In the
    proposal we stated that for each time measurement we would perform the relevant operation 1000
    times and then take the average of that.  In practice, this made the tests run for longer than
    we could afford. Because the results given by performing 100 runs were comparable, we settled for that.  
    We expect the general result to be similar; refer to section\footnote{ref}for details
    about what \emph{could} have gone wrong.

    \section{Results}

    Our results were great!  They are best analysed with a two-sample independent T test.  We had an
    equal number of measurements about group 1 (with mean $x_1$ and standard deviation $\sigma_1$ and
    about group 2 (with mean $x_2$ and standard deviation $\sigma_2$).  From this we can conclude that
    data structure 1 is, on average, unquestionably superior.

    \subsection{Errors possibly caused by an imperfect implementation}

    There were a number of measurements that did not follow the underlying theory and are likely a
    consequence of the hardware involved.

    In particular, constructing a small hash table\footnote{state the number of elements in a "small" hash table}
    after constructing many big ones takes a disproportionally long time.  Only the first hash table is affected, and the time is 			comparable to the time required to construct a hash table with 4000 elements, which is illogical.
    We assume this is caused by the allocator but the reason for such behaviour is still unclear.  The fact
    that a hashtable needs contiguous memory is likely to have some influence, as well as the
    allocator returning memory to the operating system.  However, we reason the latter should also have happened
    with the other structures.

    A similar issue occurs around 4000 or 8000 elements (depending on data structure).  For all three
    datastructures, the cost of inserting and querying grows at a higher rate than before.  This is difficult to
    quantify\footnote{Can we do some sort of linear regression to prove this is the case?}, but it
    appears some extra load is added at that point.  We suspect it is related either to the
    allocator having to acquire more memory or the CPU cache running out and more cache misses occuring.
    We would measure this as well, if there was more time.

    A potentially related issue is that the standard deviation of our measurements rises around the
    same point.  That increases the likelihood of it being allocator-related, as that involves a
    system call and could cause scheduling issues\footnote{elaborate/ref?}, despite the system being 
    otherwise very close to idle.

    Also, we acquired counterintuitive results showing that inserting or querying a full path is
    faster than querying only the domain name.  This can partially be explained\footnote{Not really,
    but ehhh.}by the fact that the string implementation we used was based on reference counting.
    %(If it'd also provide SSO, this would be easy to explain, but the darn thing doesn't.)

    \section{Conclusions}
    Our main point of research was researching which associative data structure performed best on
    a various number of insertions, in terms of the speed of insertions and look-ups, and memory usage.
    In terms of speed we can state with certainty that the hash table is the superior data structure.
    It outperforms the other two in almost every category, most significantly in larger datastructures where
    it often takes less than half as much time as the skiplist and less than two thirds the time of the binary search tree.
    In most cases, the binary search tree is a bit slower, whereas the skip list is the slowest structure in
    virtually every case. The only motivation for using a hashmap for this problem appears to be in memory usage,
    although the advantage gained from this is relatively small.

    Our secondary point of research encompassed the effect of key complexity on associative data structures.
    Our different types of keys were IP adresses, domain names and full path URLs. We have seen that as expected,
    IP adresses as keys provide a significant improvement over the other two, and uses less memory as well. Interesting to
    note is that we found the only advantage of the binary search tree over the hashmap here, is that insertions of
    a small amount of IP adresses, up to around $2^13$ insertions, is faster. So when writing an application with very
    few keys of low complexity, and with more insertions than queries, the binary search tree may be the best option. Finally,
    we have found that the domain key type does not outperform the full path key type. It is not significantly better
    if it is better at all.
    %The skip list uses the least memory but is awful in every other imaginable way.  The hashmap is
    %usually the best, except when we're dealing with a small number of integer keys and doing far more
    %inserts than queries.  Note that this doesn't actually happen in practice.

    \section{Reflection}
    In general, we are content with the results of our research. We were able to provide clear and concise answers to our
    research questions. Even though our research questions may appear broad, we have defined a good scope in our research
    proposal and have virtually not altered this scope in our research. One small change is that we also included datasets of
    size $2^8$ because we saw some interesting results in very small sets. In our implementation we have made some significant
    changes because we ran into pratical problems. Some of these required us to change the allocator and generate our keys
    in a slightly different fashion, which we all documented in this paper.

    Since we have found that the hash map is the best associative data structure overal for this problem, we could do further
    research in different kind of hash maps. We might try different hash functions, different methods of collision resolution,
    and perhaps survey different kinds of concurrent hash maps.

    \bibliographystyle{alpha}

    \bibliography{paper}

\end{document}